---
title: "NHL Shot Effectiveness Model"
author: "Sasank Vishnubhatla"
date: "February 8th, 2019"
runtime: shiny
output:
  html_document:
    theme: simplex
    df_print: paged
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Last Update: `r Sys.time()`

# Libraries

Before we start, let's load a few libraries.

```{r libraries}
rm(list = ls())

options(warn = -1)

library(knitr)
library(ggplot2)
library(caret)
library(doParallel)

registerDoParallel(cores = (detectCores() - 1))
```

With our libraries loaded we can start loading our data.

# Data Loading

Let's read in our data.

```{r read_csv}
data.2015 = read.csv("data/2015.csv")
data.2016 = read.csv("data/2016.csv")
data.2017 = read.csv("data/2017.csv")
data.2018 = read.csv("data/2018.csv")
```

Now, we will only deal with regular season events. So let's remove the playoffs from our datasets.

```{r remove_playoffs}
get.regular.season = function(data) {
	subset(data, isPlayoffGame == 0)
}

season.2015 = get.regular.season(data.2015)
season.2016 = get.regular.season(data.2016)
season.2017 = get.regular.season(data.2017)
season.2018 = get.regular.season(data.2018)
```

Here is a table of all the columns we shall keep and what we shall rename them to.

| Old Column Name | New Column Name |
| --------------- | --------------- |
| `xCordAdjusted` | `x` |
| `yCordAdjusted` | `y` |
| `shotAngleAdjusted` | `angle` |
| `shotDistance` | `dist` |
| `teamCode` | `team` |
| `shotType` | `type` |

```{r get_helpful_data}
get.helpful.data = function(data) {
	data.frame(x = data$xCordAdjusted,
			   y = data$yCordAdjusted,
			   angle = data$shotAngleAdjusted,
			   dist = data$shotDistance,
			   type = data$shotType,
			   typeNum = as.numeric(data$shotType),
			   goal = data$goal,
			   team = data$teamCode)
}

# type:
# 1 -> empty
# 2 -> BACK
# 3 -> DEFL
# 4 -> SLAP
# 5 -> SNAP
# 6 -> TIP
# 7 -> WRAP
# 8 -> WRIST

analysis.2015 = get.helpful.data(season.2015)
analysis.2016 = get.helpful.data(season.2016)
analysis.2017 = get.helpful.data(season.2017)
analysis.2018 = get.helpful.data(season.2018)
```

Now, we can remove incomplete cases and create our machine learning model's giant data set.

```{r incomplete_cases}
analysis.2015 = analysis.2015[complete.cases(analysis.2015),]
analysis.2016 = analysis.2016[complete.cases(analysis.2016),]
analysis.2017 = analysis.2017[complete.cases(analysis.2017),]
analysis.all = rbind(analysis.2017, rbind(analysis.2016, analysis.2015))
analysis.all = analysis.all[complete.cases(analysis.all),]
analysis.2018 = analysis.2018[complete.cases(analysis.2018),]
```

Here's what `analysis.2018` looks like:

```{r view_analysis}
analysis.2018
```

Now, we need a few functions to help us select certain subsets of data. We'll define three functions: `get.team.data`, `get.shooter.data`, `get.goalie.data`.

```{r data_functions}
get.team.data = function(data, code) {
	subset(data, team == code)
}
```

# Calculating Statistics

We can calculate a few statistics, like goal (effective) percentage for a certain shot. Let's write a function to do that right now.

```{r goal_perc}
calculate.goal.percentage = function(data) {
	goals = sum(data$goal == 1)
	total = nrow(data)
	goals / total
}
```

So, for example, Penguins's goal percentage against slap shots would be calculated as follows:

```{r murray_slap}
penguins.2018 = get.team.data(analysis.2018, "PIT")
penguins.2018.slap = subset(penguins.2018, type == 4)
penguins.2018.slap.eff = calculate.goal.percentage(penguins.2018.slap)
```

Their goal percentage is `r penguins.2018.slap.eff`.

# Creating the Models

Using the `caret` package, we can build machine learning models to help us determine which shot type is best.

Let's get started with training our control.

```{r control}
control = trainControl(method = "repeatedcv", number = 5, repeats = 3)
```

Now we will train a few different types of models. Here is a list of the models we will train:

- Neural Network
- K Nearest Neighbors

```{r models}
model.nnet = train(goal ~ . -goal -team -type,
				   data = analysis.all,
				   method = "nnet",
				   trControl = control)
model.knn = train(goal ~ . -goal -team -type,
				  data = analysis.all,
				  method = "knn",
				  trControl = control)
```

Now our models have been made.

# Testing the Models

Let's test our models on the 2018 data. Here's what the testing data looks like:

```{r testing_data}
analysis.2018
```

Now, let's get our predictions:

```{r get_predictions}
nnet.prediction = predict(model.nnet, newdata = analysis.2018)
knn.prediction = predict(model.knn, newdata = analysis.2018)

nnet.prediction.data = data.frame(analysis.2018)
nnet.prediction.data$predict = nnet.prediction

knn.prediction.data = data.frame(analysis.2018)
knn.prediction.data$predict = knn.prediction
```

So, our Neural Network data looks like:

```{r nnet_data}
nnet.prediction.data
```

Our K-Nearest Neighbors data looks like:

```{r knn_data}
knn.prediction.data
```

# Visualizing the Predictions

Let us visualize how distance and the goal prediction are related.

```{r vis_knn_dist}
make.knn.dist.plot = function(data, primary, secondary, team) {
	name = paste(team, "Predicted Goal Probability versus Distance (KNN)", sep = " ")
	plot = ggplot(data) +
		geom_smooth(aes(x = dist, y = predict), method = "auto",
				 fill = secondary, color = primary) +
		labs(title = name,
			 x = "Distance from Net",
			 y = "Probability of Scoring") +
		theme_minimal()
	plot
}
```

Here's the same function, but for the neural network data.

```{r vis_nnet_dist}
make.nnet.dist.plot = function(data, primary, secondary, team) {
	name = paste(team, "Predicted Goal Probability versus Distance (NNet)", sep = " ")
	plot = ggplot(data) +
		geom_smooth(aes(x = dist, y = predict), method = "auto",
				 fill = secondary, color = primary) +
		labs(title = name,
			 x = "Distance from Net",
			 y = "Probability of Scoring") +
		theme_minimal()
	plot
}
```

Now let's visualize the frequency of a shot.

```{r vis_type_freq}
make.type.freq.plot = function(data, primary, secondary, team) {
  name = paste(team, "Frequency per Shot Type", seq = " ")
  plot = ggplot(data) +
    geom_bar(aes(x = type), stat = "count",
                   fill = secondary, color = primary) +
    labs(title = name, x = "Shot Type", y = "Count") +
    theme_minimal()
  plot
}
```

The next graph is comparing type and goal prediction.

```{r vis_knn_type}
make.knn.type.plot = function(data, primary, secondary, team) {
  name = paste(team, "Predicted Goal versus Shot Type (KNN)", seq = " ")
  plot = ggplot(data) +
    geom_jitter(aes(x = type, y = predict), fill = primary, color = secondary) +
    labs(title = name, x = "Shot Type", y = "Probability of Scoring") +
    theme_minimal()
  plot
}
```

Here is the neural net flavor of the last function.

```{r vis_nnet_type}
make.nnet.type.plot = function(data, primary, secondary, team) {
  name = paste(team, "Predicted Goal versus Shot Type (NNet)", seq = " ")
  plot = ggplot(data) +
    geom_jitter(aes(x = type, y = predict), fill = primary, color = secondary) +
    labs(title = name, x = "Shot Type", y = "Probability of Scoring") +
    theme_minimal()
  plot
}
```

# Examples {.tabset .tabset-fade}

## Pittsburgh Penguins

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r pit_data}
pit.nnet = get.team.data(nnet.prediction.data, "PIT")
pit.knn = get.team.data(knn.prediction.data, "PIT")
```

Let's just visualize our shots type data.

```{r pit_type_freq}
pit.type.freq = make.type.freq.plot(pit.nnet, "#000000", "#FCB514", "PIT")

pit.type.freq
```

Here is how the Penguins did with shots versus goal probability from the KNN model.

```{r pit_knn_dist}
pit.knn.dist = make.knn.dist.plot(pit.knn, "#000000", "#FCB514", "PIT")

pit.knn.dist
```

And now with the neural network.

```{r pit_nnet_dist}
pit.nnet.dist = make.nnet.dist.plot(pit.knn, "#000000", "#FCB514", "PIT")

pit.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r pit_knn_type}
pit.knn.type = make.knn.type.plot(pit.knn, "#000000", "#FCB514", "PIT")

pit.knn.type
```

And now with the neural network.

```{r pit_nnet_type}
pit.nnet.type = make.nnet.type.plot(pit.knn, "#000000", "#FCB514", "PIT")

pit.nnet.type
```

## Vegas Golden Knights

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r vgk_data}
vgk.nnet = get.team.data(nnet.prediction.data, "VGK")
vgk.knn = get.team.data(knn.prediction.data, "VGK")
```

Let's just visualize our shots type data.

```{r vgk_type_freq}
vgk.type.freq = make.type.freq.plot(vgk.nnet, "#B4975A", "#333F42", "VGK")

vgk.type.freq
```

Here is how the Golden Knights did with shots versus goal probability from the KNN model.

```{r vgk_knn_dist}
vgk.knn.dist = make.knn.dist.plot(vgk.knn, "#B4975A", "#333F42", "VGK")

vgk.knn.dist
```

And now with the neural network.

```{r vgk_nnet_dist}
vgk.nnet.dist = make.nnet.dist.plot(vgk.knn, "#B4975A", "#333F42", "VGK")

vgk.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r vgk_knn_type}
vgk.knn.type = make.knn.type.plot(vgk.knn, "#B4975A", "#333F42", "VGK")

vgk.knn.type
```

And now with the neural network.

```{r vgk_nnet_type}
vgk.nnet.type = make.nnet.type.plot(vgk.knn, "#B4975A", "#333F42", "VGK")

vgk.nnet.type
```

## Winnipeg Jets

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r wpg_data}
wpg.nnet = get.team.data(nnet.prediction.data, "WPG")
wpg.knn = get.team.data(knn.prediction.data, "WPG")
```

Let's just visualize our shots type data.

```{r wpg_type_freq}
wpg.type.freq = make.type.freq.plot(wpg.nnet, "#041E42", "#004C97", "WPG")

wpg.type.freq
```

Here is how the Jets did with shots versus goal probability from the KNN model.

```{r wpg_knn_dist}
wpg.knn.dist = make.knn.dist.plot(wpg.knn, "#041E42", "#004C97", "WPG")

wpg.knn.dist
```

And now with the neural network.

```{r wpg_nnet_dist}
wpg.nnet.dist = make.nnet.dist.plot(wpg.knn, "#041E42", "#004C97", "WPG")

wpg.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r wpg_knn_type}
wpg.knn.type = make.knn.type.plot(wpg.knn, "#041E42", "#004C97", "WPG")

wpg.knn.type
```

And now with the neural network.

```{r wpg_nnet_type}
wpg.nnet.type = make.nnet.type.plot(wpg.knn, "#041E42", "#004C97", "WPG")

wpg.nnet.type
```

## Tampa Bay Lightning

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r tbl_data}
tbl.nnet = get.team.data(nnet.prediction.data, "T.B")
tbl.knn = get.team.data(knn.prediction.data, "T.B")
```

Let's just visualize our shots type data.

```{r tbl_type_freq}
tbl.type.freq = make.type.freq.plot(tbl.nnet, "#002868", "#AAB1BF", "T.B")

tbl.type.freq
```

Here is how the Lightning did with shots versus goal probability from the KNN model.

```{r tbl_knn_dist}
tbl.knn.dist = make.knn.dist.plot(tbl.knn, "#002868", "#AAB1BF", "T.B")

tbl.knn.dist
```

And now with the neural network.

```{r tbl_nnet_dist}
tbl.nnet.dist = make.nnet.dist.plot(tbl.knn, "#002868", "#AAB1BF", "T.B")

tbl.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r tbl_knn_type}
tbl.knn.type = make.knn.type.plot(tbl.knn, "#002868", "#AAB1BF", "T.B")

tbl.knn.type
```

And now with the neural network.

```{r tbl_nnet_type}
tbl.nnet.type = make.nnet.type.plot(tbl.knn, "#002868", "#AAB1BF", "T.B")

tbl.nnet.type
```

## Toronto Maple Leafs

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r tor_data}
tor.nnet = get.team.data(nnet.prediction.data, "TOR")
tor.knn = get.team.data(knn.prediction.data, "TOR")
```

Let's just visualize our shots type data.

```{r tor_type_freq}
tor.type.freq = make.type.freq.plot(tor.nnet, "#003E7E", "#FFFFFF", "TOR")

tor.type.freq
```

Here is how the Maple Leafs did with shots versus goal probability from the KNN model.

```{r tor_knn_dist}
tor.knn.dist = make.knn.dist.plot(tor.knn, "#003E7E", "#FFFFFF", "TOR")

tor.knn.dist
```

And now with the neural network.

```{r tor_nnet_dist}
tor.nnet.dist = make.nnet.dist.plot(tor.knn, "#003E7E", "#FFFFFF", "TOR")

tor.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r tor_knn_type}
tor.knn.type = make.knn.type.plot(tor.knn, "#003E7E", "#FFFFFF", "TOR")

tor.knn.type
```

And now with the neural network.

```{r tor_nnet_type}
tor.nnet.type = make.nnet.type.plot(tor.knn, "#003E7E", "#FFFFFF", "TOR")

tor.nnet.type
```

## Colorado Avalance

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r col_data}
col.nnet = get.team.data(nnet.prediction.data, "COL")
col.knn = get.team.data(knn.prediction.data, "COL")
```

Let's just visualize our shots type data.

```{r col_type_freq}
col.type.freq = make.type.freq.plot(col.nnet, "#6F263D", "#236192", "COL")

col.type.freq
```

Here is how the Avalance did with shots versus goal probability from the KNN model.

```{r col_knn_dist}
col.knn.dist = make.knn.dist.plot(col.knn, "#6F263D", "#236192", "COL")

col.knn.dist
```

And now with the neural network.

```{r col_nnet_dist}
col.nnet.dist = make.nnet.dist.plot(col.knn, "#6F263D", "#236192", "COL")

col.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r col_knn_type}
col.knn.type = make.knn.type.plot(col.knn, "#6F263D", "#236192", "COL")

col.knn.type
```

And now with the neural network.

```{r col_nnet_type}
col.nnet.type = make.nnet.type.plot(col.knn, "#6F263D", "#236192", "COL")

col.nnet.type
```

## Calgary Flames

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r cgy_data}
cgy.nnet = get.team.data(nnet.prediction.data, "CGY")
cgy.knn = get.team.data(knn.prediction.data, "CGY")
```

Let's just visualize our shots type data.

```{r cgy_type_freq}
cgy.type.freq = make.type.freq.plot(cgy.nnet, "#C8102E", "#F1BE48", "CGY")

cgy.type.freq
```

Here is how the Flames did with shots versus goal probability from the KNN model.

```{r cgy_knn_dist}
cgy.knn.dist = make.knn.dist.plot(cgy.knn, "#C8102E", "#F1BE48", "CGY")

cgy.knn.dist
```

And now with the neural network.

```{r cgy_nnet_dist}
cgy.nnet.dist = make.nnet.dist.plot(cgy.knn, "#C8102E", "#F1BE48", "CGY")

cgy.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r cgy_knn_type}
cgy.knn.type = make.knn.type.plot(cgy.knn, "#C8102E", "#F1BE48", "CGY")

cgy.knn.type
```

And now with the neural network.

```{r cgy_nnet_type}
cgy.nnet.type = make.nnet.type.plot(cgy.knn, "#C8102E", "#F1BE48", "CGY")

cgy.nnet.type
```

## New York Islanders

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r nyi_data}
nyi.nnet = get.team.data(nnet.prediction.data, "NYI")
nyi.knn = get.team.data(knn.prediction.data, "NYI")
```

Let's just visualize our shots type data.

```{r nyi_type_freq}
nyi.type.freq = make.type.freq.plot(nyi.nnet, "#00539B", "#F47D30", "NYI")

nyi.type.freq
```

Here is how the Islanders did with shots versus goal probability from the KNN model.

```{r nyi_knn_dist}
nyi.knn.dist = make.knn.dist.plot(nyi.knn, "#00539B", "#F47D30", "NYI")

nyi.knn.dist
```

And now with the neural network.

```{r nyi_nnet_dist}
nyi.nnet.dist = make.nnet.dist.plot(nyi.knn, "#00539B", "#F47D30", "NYI")

nyi.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r nyi_knn_type}
nyi.knn.type = make.knn.type.plot(nyi.knn, "#00539B", "#F47D30", "NYI")

nyi.knn.type
```

And now with the neural network.

```{r nyi_nnet_type}
nyi.nnet.type = make.nnet.type.plot(nyi.knn, "#00539B", "#F47D30", "NYI")

nyi.nnet.type
```

## Washington Capitals

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r wsh_data}
wsh.nnet = get.team.data(nnet.prediction.data, "WSH")
wsh.knn = get.team.data(knn.prediction.data, "WSH")
```

Let's just visualize our shots type data.

```{r wsh_type_freq}
wsh.type.freq = make.type.freq.plot(wsh.nnet, "#041E42", "#C8102E", "WSH")

wsh.type.freq
```

Here is how the Cawshals did with shots versus goal probability from the KNN model.

```{r wsh_knn_dist}
wsh.knn.dist = make.knn.dist.plot(wsh.knn, "#041E42", "#C8102E", "WSH")

wsh.knn.dist
```

And now with the neural network.

```{r wsh_nnet_dist}
wsh.nnet.dist = make.nnet.dist.plot(wsh.knn, "#041E42", "#C8102E", "WSH")

wsh.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r wsh_knn_type}
wsh.knn.type = make.knn.type.plot(wsh.knn, "#041E42", "#C8102E", "WSH")

wsh.knn.type
```

And now with the neural network.

```{r wsh_nnet_type}
wsh.nnet.type = make.nnet.type.plot(wsh.knn, "#041E42", "#C8102E", "WSH")

wsh.nnet.type
```

## Philadelphia Flyers

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r phi_data}
phi.nnet = get.team.data(nnet.prediction.data, "PHI")
phi.knn = get.team.data(knn.prediction.data, "PHI")
```

Let's just visualize our shots type data.

```{r phi_type_freq}
phi.type.freq = make.type.freq.plot(phi.nnet, "#F74902", "#000000", "PHI")

phi.type.freq
```

Here is how the Flyers did with shots versus goal probability from the KNN model.

```{r phi_knn_dist}
phi.knn.dist = make.knn.dist.plot(phi.knn, "#F74902", "#000000", "PHI")

phi.knn.dist
```

And now with the neural network.

```{r phi_nnet_dist}
phi.nnet.dist = make.nnet.dist.plot(phi.knn, "#F74902", "#000000", "PHI")

phi.nnet.dist
```

Now let's look at shot type versus goal probability.

```{r phi_knn_type}
phi.knn.type = make.knn.type.plot(phi.knn, "#F74902", "#000000", "PHI")

phi.knn.type
```

And now with the neural network.

```{r phi_nnet_type}
phi.nnet.type = make.nnet.type.plot(phi.knn, "#F74902", "#000000", "PHI")

phi.nnet.type
```

## Boston Bruins

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r bos_data}
bos.nnet = get.team.data(nnet.prediction.data, "BOS")
bos.knn = get.team.data(knn.prediction.data, "BOS")
```

Let's just visualize our shots type data.

```{r bos_type_freq}
bos.type.freq = make.type.freq.plot(bos.nnet, "#FFB81C", "#000000", "BOS")

bos.type.freq
```

Here is how the Bruins did with shots versus goal probability from the KNN model.

```{r bos_knn_dist}
bos.knn.dist = make.knn.dist.plot(bos.knn, "#FFB81C", "#000000", "BOS")

bos.knn.dist
```

And now with the neural network.

```{r bos_nnet_dist}
bos.nnet.dist = make.nnet.dist.plot(bos.knn, "#FFB81C", "#000000", "BOS")

bos.nnet.dist
```

Now let's look at shot type versus goal probability.


```{r bos_knn_type}
bos.knn.type = make.knn.type.plot(bos.knn, "#FFB81C", "#000000", "BOS")

bos.knn.type
```

And now with the neural network.

```{r bos_nnet_type}
bos.nnet.type = make.nnet.type.plot(bos.knn, "#FFB81C", "#000000", "BOS")

bos.nnet.type
```

## San Jose Sharks

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r sjs_data}
sjs.nnet = get.team.data(nnet.prediction.data, "S.J")
sjs.knn = get.team.data(knn.prediction.data, "S.J")
```

Let's just visualize our shots type data.

```{r sjs_type_freq}
sjs.type.freq = make.type.freq.plot(sjs.nnet, "#006D75", "#EA7200", "S.J")

sjs.type.freq
```

Here is how the Sharks did with shots versus goal probability from the KNN model.

```{r sjs_knn_dist}
sjs.knn.dist = make.knn.dist.plot(sjs.knn, "#006D75", "#EA7200", "S.J")

sjs.knn.dist
```

And now with the neural network.

```{r sjs_nnet_dist}
sjs.nnet.dist = make.nnet.dist.plot(sjs.knn, "#006D75", "#EA7200", "S.J")

sjs.nnet.dist
```

Now let's look at shot type versus goal probability.


```{r sjs_knn_type}
sjs.knn.type = make.knn.type.plot(sjs.knn, "#006D75", "#EA7200", "S.J")

sjs.knn.type
```

And now with the neural network.

```{r sjs_nnet_type}
sjs.nnet.type = make.nnet.type.plot(sjs.knn, "#006D75", "#EA7200", "S.J")

sjs.nnet.type
```

## Dallas Stars

Let's get our data from the Neural Network and K-Nearest Neighbors algorithms.

```{r dal_data}
dal.nnet = get.team.data(nnet.prediction.data, "DAL")
dal.knn = get.team.data(knn.prediction.data, "DAL")
```

Let's just visualize our shots type data.

```{r dal_type_freq}
dal.type.freq = make.type.freq.plot(dal.nnet, "#006847", "#8F8F8C", "DAL")

dal.type.freq
```

Here is how the Stars did with shots versus goal probability from the KNN model.

```{r dal_knn_dist}
dal.knn.dist = make.knn.dist.plot(dal.knn, "#006847", "#8F8F8C", "DAL")

dal.knn.dist
```

And now with the neural network.

```{r dal_nnet_dist}
dal.nnet.dist = make.nnet.dist.plot(dal.knn, "#006847", "#8F8F8C", "DAL")

dal.nnet.dist
```

Now let's look at shot type versus goal probability.


```{r dal_knn_type}
dal.knn.type = make.knn.type.plot(dal.knn, "#006847", "#8F8F8C", "DAL")

dal.knn.type
```

And now with the neural network.

```{r dal_nnet_type}
dal.nnet.type = make.nnet.type.plot(dal.knn, "#006847", "#8F8F8C", "DAL")

dal.nnet.type
```